---
title: "PA1 Template"
subtitle: "Reproducible Research Course"
author: "Carlos Gabriel Guerra Farfán"
date: "10/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the Data

The data for this assignment can be downloaded from the course web site: Dataset: Activity monitoring data [52K] The variables included in this dataset are: steps: Number of steps taking in a 5-minute interval (missing values are coded as NA) date: The date on which the measurement was taken in YYYY-MM-DD format interval: Identifier for the 5-minute interval in which measurement was taken The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

```{r loaddata}
# Load data
if (!file.exists("activity.csv") )
    {
     dlurl <- 'http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
     download.file(dlurl,destfile='repdata%2Fdata%2Factivity.zip',mode='wb')  
     unzip('repdata%2Fdata%2Factivity.zip')
    }

# Read data
data <- read.csv("activity.csv")
```

## 3a What is mean total number of steps taken per day?

```{r mean_median}
steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="green",xlab="Number of Steps")
rmean <- mean(steps_by_day$steps)
rmean
rmedian <- median(steps_by_day$steps)
rmedian
```

## 3b What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r pressure, echo=FALSE}
steps_by_interval <- aggregate(steps ~ interval, data, mean)
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r mean}
max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
max_interval
```

## 3c Imputing missing values
### 1 Calculate and report the total number of missing values in the dataset

```{r na_sum}
NATotal <- sum(!complete.cases(data))
NATotal
```

### 2 Using Mean for the day compute missing values

```{r fill_missing_value}
StepsAverage <- aggregate(steps ~ interval, data = data, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(data)) {
    obs <- data[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}
```

### 3 Create a new dataset including the imputed missing values

```{r }
new_activity <- data
new_activity$steps <- fillNA
```

## 4 Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r getpatternsbyweekdays}
StepsTotalUnion <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
hist(StepsTotalUnion$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
#Create Histogram to show difference. 
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "green"), lwd=10)
#Calculate Mean
rmeantotal <- mean(StepsTotalUnion$steps)
rmeantotal
#Calculate Median
rmediantotal <- median(StepsTotalUnion$steps)
rmediantotal
#Do these values differ from the estimates from the first part of the assignment?
rmediandiff <- rmediantotal - rmedian
rmediandiff
rmeandiff <- rmeantotal - rmean
rmeandiff
```

### 3d Are there differences in activity patterns between weekdays and weekends?

```{r  }
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
new_activity$dow = as.factor(ifelse(is.element(weekdays(as.Date(new_activity$date)),weekdays), "Weekday", "Weekend"))
StepsTotalUnion <- aggregate(steps ~ interval + dow, new_activity, mean)
library(lattice)
xyplot(StepsTotalUnion$steps ~ StepsTotalUnion$interval|StepsTotalUnion$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```
